Assessing Tor's Usability as a Censorship Circumvention Tool

### 2015-05-12 cs261n-paper/paper.tex
First quote is P1.
rough ux results section

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot Tor usability study to test the usability of Tor as an anonymity tool,
targeting a specific demographic of users known to use Tor (journalists who have reported in other
countries) and aiming to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, we refer you to Appendix~\ref{sec:pilot}.

@@

Each session of 30 participants will be placed in the same simulated censorship environment. 
We start off the experiment by informing our participants that they are in an adversarial setting, 
where some websites are blocked while other websites are not. We will instruct them to visit a censored
website and a non-censored website to illustrate that their devices and the network is fully functional and
operating as expected, but that they are in a censorship environment where some websites will fail to load. 
We will take a moment to explain what Tor is, how it is generally used in this case, and ask them to complete 
the aforementioned set of tasks. To complete all the tasks, the participants will ultimately need to get the correct
configuration settings for the country they are simulated to be in.  Figure~\ref{dialog} shows one of the 
Tor configuration dialog screens that a user will need to navigate.

@@

The survey consists of 17 questions total, with a basic breakdown as follows. 
The complete list of survey questions can be found in Appendix~\ref{sec:survey}.

@@

\subsection{Usability for Security Software}
\indent \indent 
We evaluate the usability of the configuration window's interface. Generally, usability 
refers to the learnability, efficiency, memorability, error rate/recoverability, and satisfaction
during use. To achieve these goals, the HCI community has established heuristics for user
interface design: visibility of the system status, using familiar
language, preventing errors, having consistent visuals, flexibility and efficiency of use, 
error recovery, and leveraging recognition over memorization~\cite{nielsen199510}. Note
that these are not specific usability guidelines, but heuristics.

Usability has sligtly different meanings depending on the context. Some emphasize
learnability, flexibility, and ease of user. In a security context, the priorities are whatever is 
required for the security to be used effectively. As stated by Whitten \etal, a security software is 
usable if people who are expected to use it are reliably made aware of the security tasks to perform, 
are able to figure out how to successfully perform those tasks, do not make dangerous errors, and
are sufficiently comfortable with the software interface~\cite{whitten1999johnny}. Since usable security 
has user interface design goals than general consumer software, it likewise requires usability evaluation
methods which are suitable to test these properties. Our goals are to test that the Tor Browser, 
communicates an accurate conceptual model of the task to the users as quickly as possible, provide guidance to 
users at the right time, and use warnings and other interactive tools to prevent any mistakes.

\subsection{Usability and Tor}
\indent \indent  Work by Clark \etal to explore ways to simplify the Tor user interface
has led to great usability improvements~\cite{clark2007usability}.  Clark analyzed the challenges 
of using multiple Tor-related tools including TorPark, Vidalia, FoxyProxy, and TorButton to find that
none are fully satisfactory from a usability perspective. Researchers provide guidelines on 
how to incorporate the best aspects of each tool, while providing a set of guidelines drawn from
usable security and computer human interaction. Key takeaways from this work includes 
finding out that exposure to how Tor works may be difficult for users to understand, and that 
familiar, non-technical language is critical for communication. We plan to draw from Clark's
guidelines as well as other works in usable security and computer human interaction to suggest
any improvements to the Tor Browser.

This will be the first experiment to evaluate Tor's usability as a censorship circumvention tool.
Previous work on evaluating the usability of Tor as an anonymity system have been fruitful. 
Norcie \etal identified ``stop points,'' or barriers to installing and using Tor to help streamline
the interface. Key stop points included the ability to discriminate between anonymized and non-
anonymized browser windows while using tor, unclear download and install processes, and 
confusion as to why the Tor behaves in peculiar ways~\cite{norcie2012eliminating}. A follow up
study by Norcie \etal verifies that the changes recommended to the Tor Project were indeed 
effective at reducing stopping points mentioned in the previous work. The authors also present
design heuristics for designing usable anonymous systems, such as easy installation, communicating
tradeoffs made to users, and informing why, not how, precautions are taken for security guarantees~\cite{norcie2014johnny}.

@@

To first ensure that our users knew what a cognitive walkthrough is, we explain what a cognitive
walkthrough and give a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then ask our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough along the process of completing the task. We give our 
participants feedback on how to properly conduct a cognitive walkthrough~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.

@@

\subsection{Results}
\indent \indent Because of the limited size of the study, it's not possible to state with confidence what 
fraction of users will encounter major usability obstacles. However, it was effective at discovering 
and demonstrating issues that are likely to cause problems for many users. 
People did have difficulty with installing Tor Browser (principally because of the Gatekeeper 
code-signing feature on OS X), did not understand what many of the toolbar options meant, 
and were confused by most Tor-specific behaviors. For specific problems 
encountered by an individual user, we refer you to our resources below.

{\color {red}

We saw an SEO poisoning opportunity.

\begin{quote}{\it
`` I'm going to use Chrome, because that's what I use at home...
I'm just going to open it up, and I'm going to type in "download tor" because 
that's what I'm supposed to do. And I'm going to click on the first one, because,
at the top, which is `Download Tor'.''} -P
\end{quote}

\begin{quote}{\it
``Okay. So I just would type in `tor' in the uh search engine and `Tor Project'...
`Anonymity Online'... I guess this is it. And I click. And the there is a "Download Tor", 
so I am going to click this.''} -P2
\end{quote}

@@

@@

Something unexpected which interfered with Tor's usability is that Mac's usable 
security feature which prevents opening unsigned code made Tor much less usable.

@@

Lessons learned: 
Leverage visuals and interface design to prioritize communication to users. Most of the 
important information was communicated with paragraphs of text that users actively 
communicated that they did not read or did not want to read and only read because of 
issues they were experiencing.

@@

Most participants did not have any interest or comprehension for advanced features.
In usability literature, it is good to just configure these for users, but Tor does not track
information on users. Maybe it's good to have failsafe defaults and less options.

@@

\begin{quote}{\it
``So yeah, there was an onion. Small green onion. And yeah, I guess, it's new identity.
So I'm guessing it's the same thing as with Google, when you open a page and you don't 
want the history to show up or something. And then you can, it's like you're disconnected 
from your Gmail account and everything, you can type that as a different user.''} -P3 
\end{quote}
}

>>>>

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot Tor usability study to test the usability of Tor as an anonymity tool,
targeting a specific demographic of users known to use Tor (journalists who have reported in other
countries) and aiming to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, we refer you to Appendix ~\ref{sec:pilot}.

@@

Each session of 30 participants will be placed in the same simulated censorship environment. 
We start off the experiment by informing our participants that they are in an adversarial setting, 
where some websites are blocked while other websites are not. We will instruct them to visit a censored
website and a non-censored website to illustrate that their devices and the network is fully functional and
operating as expected, but that they are in a censorship environment where some websites will fail to load. 
We will take a moment to explain what Tor is, how it is generally used in this case, and ask them to complete 
the aforementioned set of tasks. To complete all the tasks, the participants will ultimately need to get the correct
configuration settings for the country they are simulated to be in.  Figure ~\ref{dialog} shows one of the 
Tor configuration dialog screens that a user will need to navigate.

@@

The survey consists of 17 questions total, with a basic breakdown as follows. 
The complete list of survey questions can be found in Appendix ~\ref{sec:survey}.

@@

\subsection{Usability for Security Software}
\indent \indent 
We evaluate the usability of the configuration window's interface. Generally, usability 
refers to the learnability, efficiency, memorability, error rate/recoverability, and satisfaction
during use. To achieve these goals, the HCI community has established heuristics for user
interface design: visibility of the system status, using familiar
language, preventing errors, having consistent visuals, flexibility and efficiency of use, 
error recovery, and leveraging recognition over memorization ~\cite{nielsen199510}. Note
that these are not specific usability guidelines, but heuristics.

Usability has sligtly different meanings depending on the context. Some emphasize
learnability, flexibility, and ease of user. In a security context, the priorities are whatever is 
required for the security to be used effectively. As stated by Whitten \etal, a security software is 
usable if people who are expected to use it are reliably made aware of the security tasks to perform, 
are able to figure out how to successfully perform those tasks, do not make dangerous errors, and
are sufficiently comfortable with the software interface ~\cite{whitten1999johnny}. Since usable security 
has user interface design goals than general consumer software, it likewise requires usability evaluation
methods which are suitable to test these properties. Our goals are to test that the Tor Browser, 
communicates an accurate conceptual model of the task to the users as quickly as possible, provide guidance to 
users at the right time, and use warnings and other interactive tools to prevent any mistakes.

\subsection{Usability and Tor}
\indent \indent  Work by Clark \etal to explore ways to simplify the Tor user interface
has led to great usability improvements ~\cite{clark2007usability}.  Clark analyzed the challenges 
of using multiple Tor-related tools including TorPark, Vidalia, FoxyProxy, and TorButton to find that
none are fully satisfactory from a usability perspective. Researchers provide guidelines on 
how to incorporate the best aspects of each tool, while providing a set of guidelines drawn from
usable security and computer human interaction. Key takeaways from this work includes 
finding out that exposure to how Tor works may be difficult for users to understand, and that 
familiar, non-technical language is critical for communication. We plan to draw from Clark's
guidelines as well as other works in usable security and computer human interaction to suggest
any improvements to the Tor Browser.

This will be the first experiment to evaluate Tor's usability as a censorship circumvention tool.
Previous work on evaluating the usability of Tor as an anonymity system have been fruitful. 
Norcie \etal identified ``stop points,'' or barriers to installing and using Tor to help streamline
the interface. Key stop points included the ability to discriminate between anonymized and non-
anonymized browser windows while using tor, unclear download and install processes, and 
confusion as to why the Tor behaves in peculiar ways ~\cite{norcie2012eliminating}. A follow up
study by Norcie \etal verifies that the changes recommended to the Tor Project were indeed 
effective at reducing stopping points mentioned in the previous work. The authors also present
design heuristics for designing usable anonymous systems, such as easy installation, communicating
tradeoffs made to users, and informing why, not how, precautions are taken for security guarantees ~\cite{norcie2014johnny}.

@@

To first ensure that our users knew what a cognitive walkthrough is, we explain what a cognitive
walkthrough and give a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then ask our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough along the process of completing the task. We give our 
participants feedback on how to properly conduct a cognitive walkthrough ~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.

@@

\subsection{Results}
\indent \indent This experiment uncovered a powerful vector for attack, an unlikely source
of usability issues, and insight into how to better communicate to users so that they can be 
empowered to use Tor Browser and its advanced settings. 
Because of the limited size of the study, it's not possible to state with confidence what 
fraction of users will encounter major usability obstacles. However, it was effective at discovering 
and demonstrating issues that are likely to cause problems for many users. 
People did have difficulty with installing Tor Browser (principally because of the Gatekeeper 
code-signing feature on OS X), did not understand what many of the toolbar options meant, 
and were confused by most Tor-specific behaviors. For specific problems 
encountered by an individual user, we refer you to our resources online.

\subsubsection{How to Attack Users}
During the task completion process, we observed that all our participants are likely 
susceptible to a search engine optimization (SEO) poisoning attack. Most of our 
participants  were on autopilot to complete this standard download task and 
were also not aware of the possibility that their download may come from an 
illegitimate source. Usually, our participants picked whichever was the first search
result and believed it to be a good enough source for the download:

\begin{quote}{\it
`` I'm going to use Chrome, because that's what I use at home...
I'm just going to open it up, and I'm going to type in "download tor" because 
that's what I'm supposed to do. And I'm going to click on the first one, because,
at the top, which is `Download Tor'.''} -P1
\end{quote}

@@

There was one participant who did not pick the first search result, but to save themselves
time while completing the task:

@@

Luckily, all of our users managed to download Tor browser from the legitimate source. However, 
upon searching through various search engines with a clear history, we found that cnet.com or techsource.com 
 appears as the first search result. Defenses against SEO optimization would be prudent to employ.

\subsubsection{Fighting Mac's Security Features}
The biggest usability struggle was the result of tensions between Mac's security feature 
and Tor's usability, rather than a specific feature of the Tor Browser itself.

@@

\subsubsection{On Visual Communication}
Through observing our participants, 
we believe visual elements into their user interface design may be the key to 
communicating with users.

@@

\subsubsection{User Comprehension} 
Most of our participants did not have the expertise to understand 
the advanced setting for Tor Browser, and were wary to user it, but explaining why an option 
tis available to the user may increase comprehension and tailored use.

@@

\begin{quote}{\it
``So yeah, there was an onion. Small green onion. And yeah, I guess, it's new identity.
So I'm guessing it's the same thing as with Google, when you open a page and you don't 
want the history to show up or something. And then you can, it's like you're disconnected 
from your Gmail account and everything, you can type that as a different user.''} -P3 
\end{quote}
